Drought vulnerability of tree growth

### 2024-07-10 scripts/explore_analysis.Rmd
update exploratory analysis script

dendro.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  filter(is.na(DBH1_flag), DBH1_error=="none",
         win.decimal.error=="none", is.na(win.flag))%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(calcDBH=calcDBH[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   Cii=max(Cii, na.rm=T))

dbh.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  filter(is.na(dbh.flag), dbh.decimal.error=="none")%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(dbh=DbhDendro[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   Cii=max(Cii, na.rm=T))

@@

#find the StemTags of trees that are in the dendro data
#match using the DBH04 column

#identify multistem trees from the census
multistem<-hkk_census %>%
  group_by(tag)%>%
  dplyr::summarise(n.stem=n())%>%
  rename(Tag=tag)

multistem<-merge(multistem, trees, by="Tag", all.x=T)

multistem<-multistem %>%
rename(tag=Tag)%>%
dplyr::mutate(dbh= (as.numeric(DBH.04)*10))

hkk.stem3.dendro<-hkk.stem3 %>%
  filter(tag %in% multistem$tag)

#add the StemTag values from hkk.stem3.dendro to multistem 
#match based on matching dbh and tag

multistem<- multistem %>%
dplyr::inner_join(select(hkk.stem3.dendro, c("tag", "StemTag", "dbh")), 
by=c("tag", "dbh"))

sum(multistem$tag != multistem$StemTag)
#there are no stemtags with values different from Tag
#all stems with dendrobands have them on the main stem?

#hkk_census<-hkk_census %>%
#  filter(StemTag %in% multistem$StemTag)

@@

#plot average increment for the top 4 species

hkk_census2<-hkk_census %>%
dplyr::rename(Tag=tag)

@@

#add long-term increment to dendro data
dendro.data<-merge(dendro.data, hkk_census, by="Tag", all.x=T)

@@

dendro_inc<- dendro.data %>%
  filter(Cno %in% alt_census)%>%
  arrange(Cno)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, dplyr::lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-dplyr::lag(Cno, n=1, default=NA, order_by=Tag),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

dbh_inc<- dbh.data %>%
  filter(Cno %in% alt_census)%>%
  arrange(Cno)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(dbh_min1=dplyr::lag(dbh, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, dplyr::lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=dbh-dbh_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-dplyr::lag(Cno, n=1, default=NA, order_by=Tag),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

```
```{r species wise increments, echo=F}

#filter increments for two species

saccli<-dendro_inc %>%
  filter(sp == "SACCLI")

hopeod <- dendro_inc %>%
  filter(sp == "HOPEOD")

tetrnu<-dendro_inc %>%
  filter(sp == "TETRNU")

polyvi <- dendro_inc %>%
  filter(sp == "POLYVI")

#spaghetti plot saccli and hopeod increments over time

saccli_plot<-ggplot() +
  geom_line(data = saccli, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for SACCLI (deciduous)")+
  theme_bw()

hopeod_plot<-ggplot() +
  geom_line(data = hopeod, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for HOPEOD (evergreen)")+
  theme_bw()

tetrnu_plot<-ggplot() +
  geom_line(data = tetrnu, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for TETRNU (deciduous)")+
  theme_bw()

polyvi_plot<-ggplot() +
  geom_line(data = polyvi, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for POLYVI (evergreen)")+
  theme_bw()

library(gridExtra)
png("growth-precip-thailand/doc/display/ind_increments.png", width=8, height=8, units="in", res=300)
grid.arrange(hopeod_plot, tetrnu_plot, polyvi_plot, saccli_plot, nrow=2)
dev.off()

@@

top4_dbh<-dbh_inc %>%
filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI"))

@@

spagplot_top4_dbh<-ggplot() +
  #species plots
  geom_line(data=top4_dbh %>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc, 
  group=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), 
  col=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_line(data=top4_dbh %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="black", size=2)+
  scale_color_viridis_d()+
  guides(col=guide_legend(title="species"))+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species")+
  theme_bw()

@@

#individual panels using facet_wrap
spagplot_dbh_ind<-ggplot() +
  geom_line(data = top4_dbh, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=top4_dbh %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=top4_dbh %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  facet_wrap(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~., 
  nrow=2)+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species from DBH measurements")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top4_dbh_ind.png", width=8, height=8, units="in", res=300)
spagplot_dbh_ind
dev.off()

png("growth-precip-thailand/doc/display/spaghetti_top4_dbh.png", width=4, height=4, units="in", res=300)
spagplot_top4_dbh
dev.off()

@@

spagplot_top10_dbh<-ggplot()+
  #species plots
  geom_line(data=dbh_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dbh_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dbh_inc %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(median_inc=mean(median_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col=guide_legend("species"))+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 10 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top10_dbh.png", width=4, height=4, units="in", res=300)
spagplot_top10_dbh
dev.off()

@@

@@

```
```{r comparing dbh and dendro, echo=F}

#merge dbh and dendro data by Tag and Cno

inc_all<-merge(dendro_inc, dbh_inc, by=c("Tag", "Cno"), all=T)

#plot increments from dbh and dendro data

png("growth-precip-thailand/doc/display/dbh_dendro_compare.png", width=4, height=4, units="in", res=300)
ggplot(inc_all %>% filter(sp.x %in% c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), aes(x=inc_annual.x, y=inc_annual.y, 
col=factor(sp.x, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_point()+
  scale_color_viridis_d()+
  geom_abline(intercept=0, slope=1)+
  xlab("dendroband increment (cm)")+
  ylab("dbh increment (cm)")+
  guides(col=guide_legend(title="species"))+
  ggtitle("comparison of dbh and dendro increments")+
  theme_bw()
dev.off()
```

@@

resistance_2010_top4 <- top4 %>%
  filter(Cno %in% c(3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

resistance_2015_top4 <- top4 %>%
  filter(Cno %in% c(11, 13, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

@@

#map of resistance.div quantiles for top 4 species
resistance_all<-rbind(resistance_2010_top4, resistance_2015_top4)
resistance_all$yr<-c(rep(2010, nrow(resistance_2010_top4)), rep(2015, nrow(resistance_2015_top4)))

@@

resistance_2010 <- dendro_2010 %>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))

>>>>

dendro.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  filter(is.na(DBH1_flag), DBH1_error=="none",
         win.decimal.error=="none", is.na(win.flag))%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(calcDBH=calcDBH[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   n.dendro=min(n.dendro), #which dendroband is it from?
                   Cii=max(Cii, na.rm=T))

dbh.data<-dendrobyband %>%
  dplyr::rename(sp=SPCODE.UPDATE)%>%
  mutate(year=year(Date2)) %>%
  filter(is.na(dbh.flag), dbh.decimal.error=="none")%>%
  group_by(Cno, sp, Tag) %>%
  dplyr::summarise(dbh=DbhDendro[which(n.dendro==min(n.dendro))],
                   Date2=Date2[which(n.dendro==min(n.dendro))],
                   n.dendro=min(n.dendro),
                   Cii=max(Cii, na.rm=T))

@@

@@

#plot average increment for the top 4 species
hkk_census2<-hkk_census %>%
dplyr::rename(Tag=tag)

@@

#add long-term increment to dendro data
dendro.data<-merge(dendro.data, hkk_census, by="Tag", all.x=T)
dbh.data<-merge(dbh.data, hkk_census, by="Tag", all.x=T)

@@

dendro_inc<- dendro.data %>%
  filter(Cno %in% alt_census)%>%
  arrange(Cno)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
  n.dendro_diff=n.dendro-dplyr::lag(n.dendro, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, dplyr::lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-dplyr::lag(Cno, n=1, default=NA, order_by=Tag),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

dbh_inc<- dbh.data %>%
  filter(Cno %in% alt_census)%>%
  arrange(Cno)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(dbh_min1=dplyr::lag(dbh, n=1, default=NA, order_by=Tag),
  n.dendro_diff=n.dendro-dplyr::lag(n.dendro, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, dplyr::lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=dbh-dbh_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-dplyr::lag(Cno, n=1, default=NA, order_by=Tag),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)

```{r comparing dbh and dendro and identify outliers, echo=F}

#merge dbh and dendro data by Tag and Cno
inc_all<-merge(dendro_inc, dbh_inc, by=c("sp", "Tag", "Cno"), all=T)

#Perpendicular distance from point 'a' to a line with 'slope' and 'intercept'
dist_point_line <- function(a, slope, intercept) {
    b = c(1, intercept+slope)
    c = c(-intercept/slope,0)       
    v1 <- b - c
    v2 <- a - b
    m <- cbind(v1,v2)
    return(abs(det(m))/sqrt(sum(v1*v1)))
}

inc_all<-inc_all %>%
  dplyr::rename(inc_annual.dendro=inc_annual.x, 
  inc_annual.dbh=inc_annual.y,
  band.diff.dendro=n.dendro_diff.x,
  band.diff.dbh=n.dendro_diff.y) %>%
  #split into three size classes
  dplyr::mutate(size_class= ifelse(dbh<35, "small",
  ifelse(dbh<70, "medium", "large")),
  size_class=factor(size_class, levels=c("small", "medium", "large"))
  #find the deviation of the increment from the 1:1 line
  )%>%
  select(sp, Tag, Cno, inc_annual.dendro, inc_annual.dbh, dbh, 
  size_class, band.diff.dbh, band.diff.dendro)

#using the distance to the line
dev.1<-mapply(function(x)dist_point_line(c(inc_all$inc_annual.dendro[x], inc_all$inc_annual.dbh[x]), 1, 0), x=1:nrow(inc_all))
inc_all$dev.1<-dev.1

#compare dbh and dendro increments for all species

cor.all<-cor(inc_all$inc_annual.dendro, inc_all$inc_annual.dbh, use="complete.obs")

dbh_dendro_all<-ggplot(inc_all, aes(x=inc_annual.dendro, y=inc_annual.dbh, col=factor(band.diff.dendro+band.diff.dbh, levels=c(0, 1, 2))))+
  geom_point(alpha=0.5)+
  geom_abline(intercept=0, slope=1)+
  scale_color_manual(values=c("black", "orange", "tomato", "grey40"))+
  xlab("dendroband increment (cm)")+
  ylab("dbh increment (cm)")+
  geom_rug(alpha=0.3)+
  geom_text(aes(x=2, y=4, label = paste0("r = ", round(cor.all,3))), vjust = -0.5) +
  guides(col=guide_legend(title="band_diff"))+
  ggtitle("comparison of dbh and dendro increments")+
  theme_bw()

#plot distribution of deviations

pct.0.5<-round(sum(inc_all$dev.1>=0.5, na.rm=T)*100/nrow(inc_all),2)
pct.1<-round(sum(inc_all$dev.1>=1, na.rm=T)*100/nrow(inc_all),2)
pct.2<-round(sum(inc_all$dev.1>=2, na.rm=T)*100/nrow(inc_all),2)
pct.0.1<-round(sum(inc_all$dev.1>=0.1, na.rm=T)*100/nrow(inc_all),2)

png("growth-precip-thailand/doc/display/deviation_distribution.png", width=4, height=4, units="in", res=300)
ggplot()+
  geom_density(data=inc_all, aes(x=dev.1), size=1.2)+
  ggtitle("deviation of dendroband and tape-based \nincrements from each other")+
  geom_vline(xintercept=c(0.1, 0.5, 1, 2), lty=2, col="orange")+
  geom_text(aes(x=c(0.1, 0.5, 1, 2), y=5, 
  label = paste0(">",c(0.1, 0.5, 1, 2), " = ", c(pct.0.1, pct.0.5, pct.1, pct.2), "%")), 
  angle=90, vjust = 1, size=2)+
  xlab("distribution of deviations from 1:1 line")+
  theme_bw()
dev.off()
```

```{r separating outliers, echo=F}

#outlier dataframe
outliers<-inc_all %>% filter(dev.1>=0.5)
write.csv(outliers, "growth-precip-thailand/data/HKK-dendro/inc_outliers.csv", row.names=F)

#clean dataframe

dendro_inc<-merge(dendro_inc, inc_all %>% select(Tag, Cno, sp, dev.1), 
by=c("Tag", "Cno", "sp"), all.x=T)

dendro_inc_clean<-dendro_inc %>% filter(dev.1<0.5)

#plot increments from dbh and dendro data with outliers removed

cor.noouts<-dendro_inc_clean %>%
dplyr::summarise(cor=cor(inc_annual.dendro, inc_annual.dbh, use="complete.obs"))
cor.noouts<-cor.noouts$cor[1]

dbh_dendro_noouts<-ggplot(inc_all %>% filter(dev.1<0.5), aes(x=inc_annual.dendro, y=inc_annual.dbh, col=factor(band.diff.dendro+band.diff.dbh, levels=c(0, 1, 2))))+
  geom_point(alpha=0.5)+
  geom_abline(intercept=0, slope=1)+
  scale_color_manual(values=c("black", "orange", "tomato", "grey40"))+
  xlab("dendroband increment (cm)")+
  ylab("dbh increment (cm)")+
  geom_rug(alpha=0.3)+
  geom_text(aes(x=2, y=4, label = paste0("r = ", round(cor.noouts,3))), vjust = -0.5) +
  guides(col=guide_legend(title="band_diff"))+
  ggtitle("dbh and dendro increments (outliers removed)")+
  theme_bw()
dev.off()

library(gridExtra)
png("growth-precip-thailand/doc/display/dbh_dendro_compare_all.png", width=8, height=4, units="in", res=300)
grid.arrange(dbh_dendro_all, dbh_dendro_noouts, ncol=2)
dev.off()

```
```{r compare deviations by species and size, echo=F}

ggplot(inc_all %>% filter(sp.x %in% c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), aes(x=inc_annual.x, y=inc_annual.y, 
col=factor(sp.x, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_point()+
  scale_color_viridis_d()+
  geom_abline(intercept=0, slope=1)+
  xlab("dendroband increment (cm)")+
  ylab("dbh increment (cm)")+
  guides(col=guide_legend(title="species"))+
  ggtitle("comparison of dbh and dendro increments")+
  theme_bw()

#split this by size classes

cors.size<-inc_all %>% 
filter(!is.na(size_class))%>%
  group_by(size_class) %>% 
  dplyr::summarise(cor=cor(inc_annual.dendro, inc_annual.dbh, use="complete.obs"))

png("growth-precip-thailand/doc/display/dbh_dendro_compare_size.png", width=8, height=4, units="in", res=300)
ggplot(inc_all %>% filter(!is.na(size_class)), aes(x=inc_annual.dendro, y=inc_annual.dbh, 
col=size_class))+
  geom_point(alpha=0.5)+
  geom_abline(intercept=0, slope=1)+
  xlab("dendroband increment (cm)")+
  ylab("dbh increment (cm)")+
  facet_grid(.~size_class)+
  geom_text(data=cors.size, aes(x=2, y=4, label = paste0("r = ", round(cor,3))), vjust = -0.5) +
  guides(col=guide_legend(title="size class"))+
  ggtitle("comparison of dbh and dendro increments")+
  theme_bw()
dev.off()

```{r species wise increments, echo=F}

@@

top4_noouts<-dendro_inc_clean %>%
filter(sp %in% c("SACCLI", "HOPEOD", "TETRNU", "POLYVI"))

#individual panels using facet_wrap
spagplot_ind<-ggplot() +
  geom_line(data = top4, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=top4 %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=top4 %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  facet_wrap(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~., 
  nrow=2)+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top4_ind.png", width=8, height=8, units="in", res=300)
spagplot_ind
dev.off()

#individual panels using facet_wrap for the no outs dataset
spagplot_ind_noouts<-ggplot() +
  geom_line(data = top4_noouts, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=top4_noouts %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=top4_noouts %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  facet_wrap(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~., 
  nrow=2)+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species (no outliers)")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top4_ind_noouts.png", width=8, height=8, units="in", res=300)
spagplot_ind_noouts
dev.off()

@@

@@

@@

#spaghetti plot for top 10 species with outliers removed

spagplot_top10_noouts<-ggplot()+
  #species plots
  geom_line(data=dendro_inc_clean %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc_clean %>% filter(sp %in% sp_top10)%>% group_by(Cno) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc_clean %>% filter(sp %in% sp_top10)%>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(median_inc=mean(median_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col=guide_legend("species"))+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 10 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top10_noouts.png", width=4, height=4, units="in", res=300)
spagplot_top10_noouts
dev.off()

#all species with and without outliers

@@

spagplot_all_noouts<-ggplot() +
  #species plots
  geom_line(data=dendro_inc_clean %>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc_clean %>% group_by(Cno) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc_clean %>% group_by(Cno, sp) %>% 
  dplyr::summarise(median_inc=median(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(median_inc=mean(median_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=median_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col="none", group="none")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for all species (no outs)")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_all_noouts.png", width=4, height=4, units="in", res=300)
spagplot_all_noouts
dev.off()

@@

@@

#drought years 2010 and 2015
# 2010 - cno 5, 2015 - cno 15

resistance_top4 <- top4 %>%
  filter(Cno %in% c(5, 15))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(
    yr=ifelse(Cno==5, 2010, 2015),
        resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-avg_inc,
         resistance.prop = (inc_annual-avg_inc)/avg_inc)%>%
  filter(!is.infinite(resistance.div))

@@

#map of resistance.div quantiles for top 4 species

resistance_div<-ggplot(data = resistance_all, aes(x=resistance.div)) +
  geom_density(alpha=0.5) +
  facet_grid(.~yr)+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

png("growth-precip-thailand/doc/display/resistance_div_top4.png", width=6, height=4, units="in", res=300)
resistance_div
dev.off()

resistance_div_sp<-ggplot(data = resistance_all, 
aes(x=resistance.div, 
col=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")),
fill=factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_density(alpha=0.5) +
  facet_grid(factor(sp, levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))~yr)+
  scale_color_viridis_d()+ scale_fill_viridis_d()+
  guides(col="none", fill="none")+
  xlim(-5, 5)+
  xlab("inc/long-term increment")+
  geom_vline(xintercept = c(0,1), lty=2)+
  theme_bw()

png("growth-precip-thailand/doc/display/resistance_div_top4_sp.png", width=6, height=12, units="in", res=300)
resistance_div_sp
dev.off()

@@

resistance_2010 <- dendro_inc %>%
filter(Cno %in% c(3, 5))%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(inc_min1=dplyr::lag(inc_annual, n=1, default=NA, order_by=Tag),
         resistance.div = inc_annual/avg_inc,
         resistance.dif = inc_annual-inc_min1,
         resistance.prop = (inc_annual-inc_min1)/avg_inc)%>%
  filter(!is.na(inc_min1), !is.infinite(resistance.div))
